A family of accessibility measures derived from spatial interaction principles

### 2025-02-10 TRB/Musings-a-draft-pres.qmd
Edits to new text added

---
title: "Some musings Family of accessibility measures"
format: revealjs
---
```{r}
library(here)
library(knitr)
library(glue)
```

>>>>

---
title: "Some musings Family of accessibility measures"
format: revealjs
---
```{r}
library(here) # A Simpler Way to Find Your Files
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
library(glue) # Interpreted String Literals
#annotater::annotate_active_file() #Cool!
```

### 2025-02-10 paper/manuscript.qmd
Edits to new text added

In this respect, we welcome @wuUnifyingAccess2020's contribution for the way it provides a unifying framework to think about accessibility measures. By considering the key features of accessibility, namely travel-cost and the distribution of opportunities, these authors demonstrate that a majority of the concepts found scattered throughout the accessibility literature can be seen as particular cases of a general accessibility formula. One needs only to judiciously change the way travel-cost and opportunities are formulated, to derive almost any known accessibility measure.

With their research, @wuUnifyingAccess2020 have taken a considerable step towards clarifying the differences between various accessibility indicators. Still, we feel that there is room to further clarify another relevant aspect. @wuUnifyingAccess2020's treatment of competition deserves further scrutiny, since the unifying framework does not appear to accommodate many (or any) of the hugely popular floating catchment area methods. In recent research, however, @soukhovIntroducingSpatialAvailability2023 demonstrate that the introduction of a single constraint is sufficient to turn an accessibility measure of the following type (the most general in Wu and Levinson's framework) into a competitive measure of accessibility _à la_ 2-Step Floating Catchment Area approach:

$$
S_i \propto \sum_j g(O_i)f(c_{ij})
$$

In this paper we contend that accessibility research needs to reconnect with its roots in spatial interaction modelling. It is by looking to the past that we believe accessibility analysis can take a new step to the future. In particular, we argue that an important aspect of spatial interaction modelling, namely constraining the results to match empirical observations, was never properly reincorporated into accessibility analysis after these two streams of research began to diverge after the work of @hansen1959 and @wilson1971. As Wilson demonstrated, applying empirical constraints led to various spatial interaction models, and in this work we aim to show that the same can be done for accessibility measures. On the contrary, the lack of constraints in accessibility analysis seems to have contributed to some of the interpretability issues that plague it, particularly the fuzziness of insights beyond making statements of proportion such as higher-than/lower-than.

With this background in mind, the focus of the present paper is on the empirical constraints and the attendant proportionality constant missing from accessibility measures. We propose to map accessibility onto Wilson's family of spatial interactions models. To achieve this, we begin by tracing the development of accessibility from its origins in spatial interaction, beginning with @ravensteinLawsMigration1889 to the accessibility research of @hansen1959. We also provide evidence of a marked divergence of accessibility and spatial interaction modelling research after the work of @wilson1971. After this, we hark back to Wilson's spatial interaction model, and use it to derive a family of accessibility measures based on different types of constraints. We illustrate various members of this family with a simple numerical example and a real world data set. We then conclude by discussing the uses of these measure and their interpretation.

@@

The patterns of people's movement in space have been a subject of scientific inquiry for almost a century and a half. From as far back as Henry C. Carey's _Principles of Social Science_ [@careyPrinciplesSocialScience1858] a concern can be observed with the scientific study of human spatial interaction. It was in this work where Carey stated that "man [is] the molecule of society [and their interaction is subject to] the direct ratio of the mass and the inverse one of distance" [@mckeanManual1883, pp. 37-38]. This statement shows how investigations into human spatial interaction have often been explicitly coloured by the features of Newton's Law of Universal Gravitation, first posited in 1687's _Principia Mathematica_ and expressed as in @eq-phys-grav-prop.

@@

To be certain, the expression above, a proportionality, is one of the most famous in all of science. In brief, it states that the force of attraction $F$ between a pair of bodies $i$ and $j$ is directly _proportional_ to the product of their masses $M_i$ and $M_j$, and inversely _proportional_ to the square of the distance between them $D_{ij}$. Direct proportionality means that as the product of the masses increases, so does the force. Likewise, inverse proportionality means that as the distance increases, the force decreases. @eq-phys-grav-prop, however, does not quantify the magnitude of the force, for to do so an empirical constant is required to convert the proportionality into an equality, thus ensuring that values of the force $F$ in @eq-phys-grav-prop match the observed force of attraction between masses. In other words, @eq-phys-grav-prop needs to be _constrained_ using empirical data. Ultimately, the equation for the force is as seen in @eq-phys-grav, where $G$ is an empirically callibrated proportionality constant:

@@

Following Carey's _Principles_ of 1858, research into human spatial interaction continued in different contexts. In the late 1880s Ravenstein proposed some "Laws of Migration" based on his empirical analysis of migration flows in various countries [@ravensteinLawsMigration1885; @ravensteinLawsMigration1889]. In these works, Ravenstein posited 1) a directly proportional relationship between migration flows and the size of destinations (i.e., centres of commerce and industry), and 2) an inversely proportional relationship between the size of flows and the separation between origins and destinations. As with Carey, these propositions are similar to Newton's gravitational laws. Over time, other researchers discovered similar relationships. <!--Rilley does not cite Newton, nor Carey, nor Ravenstein, but I removed this curiosity as it's not central to our point here--> For example, @reilly1929methods formulated a law of retail gravitation, expressed in terms of equal attraction to competing retail destinations. Later, Zipf proposed a $\frac{P_1P_2}{D}$ hypothesis for the case of goods movement by railways [@zipfHypothesisCaseRailway1946], intercity personal movement [@zipfHypothesisIntercityMovement1946], and information [@zipfDeterminantsCirculationInformation1946]. The $\frac{P_1P_2}{D}$ hypothesis stated that the magnitude of flows was proportional to the product of the populations of settlements, and inversely proportional to the distance between them.

@@

\noindent where:

@@

\noindent where $D$ is the population density over an infinitesimal area $ds$ and $r$ is the distance to $i$. In @eq-stewart-population-potential-integral, $D\cdot ds$ gives an infinitesimal count of the population, say $dm$, and so, after discretizing space, @eq-stewart-population-potential-integral can be rewritten as:
$$
V_i = \sum_j \frac{M_j}{d_{ij}}
$${#eq-stewart-population-potential-sum}

@@

Stewart's formulation of demographic force, developed in the context of what he called "social physics" [@stewartPrinciples1947], was problematic. It had issues with inconsistent mathematical notation. <!--For example, in @stewartDemographicGravitationEvidence1948, $G$ is used as a proportionality constant p. 34 and then later as _demographic energy_ on p. 53.--> More seriously, though, Stewart's work was permeated by a view of humans as particles following physical laws, but tinted by unscientific ideas that were unadultered racism. For instance, he assumed that the molecular weight $\mu$ of the average American was one, but "presumably...much less than one....for an Australian aborigine" [p. 35]. Stewart's ideas about "social physics" soon fell out of favour among social scientists, but not before influencing the nascent field of accessibility research, as detailed next.

@@

From @stewartDemographicGravitationEvidence1948, we arrive to 1959 and Walter G. Hansen, whose work proved to be exceptionally influential in the accessibility literature [@hansen1959]. In his seminal paper, Hansen defined accessibility as "the potential of opportunities for interaction... a generalization of the population-over-distance relationship or _population potential_ concept developed by @stewartDemographicGravitationEvidence1948" (p. 73). As well as being a student of city and regional planning at Massachusetts Institute of Technology, Hansen was also an engineer with the Bureau of Roads, and preoccupied with the power of transportation to shape land uses in a very practical sense. @hansen1959 focused on @stewartDemographicGravitationEvidence1948's _population potential_ (expressed in @eq-stewart-population-potential-sum), leaving Stewart's other formulaic contributions and objectionable aspects of "social physics" behind. @hansen1959 recast @eq-stewart-population-potential-sum to reflect accessibility, a model of human behaviour useful to capture regularities in mobility patterns . In this equation, @hansen1959 replaced $M_j$ with _opportunities_ to derive an _opportunity potential_, or more accurately, a _potential of opportunities for interaction_ as follows:

@@

A modern re-writing of the above accounts for a variety of impedance functions beyond the inverse power $d^{-\beta}$:

@@

@@

Notice how, in this model, the proxy for the mass at the origin $W_i^{(1)}$ is replaced by the actual mass, as measured by the trips produced. Also, there is no longer a single system-wide proportionality constant, but rather a balancing factor specific to each origin, namely $A_i$, which Wilson found to be as follows:

@@

Notice how now the proxy for the mass at the destination $W_j^{(2)}$ is replaced by the actual mass, as measured by the trips attracted. As before, destination-specific balancing factors $B_j$ were derived by Wilson as:

@@

In this model both proxies for the masses are replaced with the known masses, that is, the trips produced by origin and the trips attracted by destination. And, now the balancing factors are:

@@

The work of @hansen1959 and @wilson1971 responded to important developments in the Anglo-American world at the time, in particular a need "to meet the dictates and needs of public policy for strategic land use and transportation planning" [@battyChronicleScientificPlanning1994]. Said dictates and needs were far from trivial. In the US alone, the Federal-Aid Highway Act of 1956 authorized the creation of the US Interstate Highway System, with a budget of more than one hundred billion dollars [@weinerUrbanTransportationPlanning2016; @mdotMnDOTJoins2007]. Spatial interaction modelling, with its ability to quantify trips, was incorporated into institutional modelling practices meant to "predict and provide", i.e., predict travel demand and supply transportation infrastructure [@kovatch1971modeling; @weinerUrbanTransportationPlanning2016]. Accessibility at the time did not quite have that power: it did not quantify trips, but rather something fuzzier: the less tangible "potential for interaction". In this way, where spatial interaction modelling became a key element of transportation planning practice, accessibility remained a somewhat more academic pursuit, and the two streams of literature only rarely connected.

@@

To illustrate this point, we conducted a bibliographic analysis of the literature that cites @hansen1959 and/or @wilson1971. We retrieved all relevant documents using the Web of Science "Cited References" functionality, and the digital object identifiers of @hansen1959 and @wilson1971. As a result of this search we identified `r n_h |> prettyNum(big.mark = ",")` documents that cite @hansen1959, `r n_w  |> prettyNum(big.mark = ",")` documents that cite @wilson1971, and `r n_b |> prettyNum(big.mark = ",")` that cite both. The earliest document in this corpus dates to `r min(corpus$PY)` and the most recent is from `r max(corpus$PY)`. The number of documents per year appears in @fig-docs-per-year, where we see the frequency of documents over a span of almost fifty years. In particular, we notice the remarkable growth in the number of papers that cite @hansen1959 compared to those that cite @wilson1971 or both.

@@

After compiling this corpus of documents, we used the {bibliometrix} package [@bibliometrix2017] to create a bibliographical coupling matrix. In this matrix two documents are coupled if they share at least one common reference, and the strength of the coupling increases with the number of references that the documents have in common. @fig-Fig1 presents the results. In this figure each symbol represents a document, and their distance on the plot represents the strength of bibliographical coupling: more strongly coupled documents, meaning those that share more items in their lists of references, are plotted closer to each other.

@@

As we argued above, the accessibility and spatial interaction modelling literature streams tended to evolve with relatively little contact over the years. This may explain the failure of the constraints/proportionality constant(s) of spatial interaction models to cross over to accessibility analysis. This is intriguing, considering that Wilson himself made an effort to connect his developments in spatial interaction modelling to accessibility, noting the similarity between the denominator in @eq-production-constrained-balancing-factor and Hansen's accessibility indicator. In fact, Wilson noted that Hansen's accessibility was the inverse of $A_i$, which offered a potential interpretation for the balancing factor (p. 10):

@@

$$
V^T_i = k\sum_j W^{(2)}_jf(c_{ij}) = \sum_j W^{(2)}_j\frac{Of(c_{ij})}{\sum_i\sum_j V^0_{ij}} = \sum_j W^{(2)}_j\frac{Of(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})} 
$$

@@

$$
\kappa = \frac{Of(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
$$

@@

In the case of SIM calibrated for accessibility, the doubly-constrained case has more limited applications than the singly-constrained case because the double constraints (both marginals) are so specific. In this case, the units of the population and opportunities must be able to be understood as the same. E.g., the number of population at $i$ that seek opportunities at each $j$ and the number of opportunities at $j$ that are saught by population at each $i$ must both be known to match.

>>>>

In this respect, @wuUnifyingAccess2020's contribution provides a unifying framework to think about accessibility measures, the most general of which they describe as $S_i$ in @eq-access-Lu-Levinson. By considering the key features of accessibility, namely travel-cost $f(c{ij})$ and the distribution of opportunities $g(O_i)$, @wuUnifyingAccess2020 demonstrate that a majority of the concepts found scattered throughout the accessibility literature can be seen as particular cases of a general accessibility formula. One needs only to judiciously change the way travel-cost and opportunities are formulated, to derive almost any known accessibility measure. In this way, @wuUnifyingAccess2020 have taken a considerable step towards clarifying the differences between various accessibility indicators.

$$
S_i \propto \sum_j g(O_j)f(c_{ij})
$${#eq-access-Lu-Levinson}

However, there is room to further clarify another relevant aspect: competition. @wuUnifyingAccess2020's unifying framework does not appear to accommodate many (or any) of the floating catchment area methods, a popular method within the competitive accessibility literature. There have also been more recent developments in the literature: for instance, @soukhovIntroducingSpatialAvailability2023 demonstrate that the introduction of a single constraint is sufficient to turn the general accessibility measure (i.e, @eq-access-Lu-Levinson) into a competitive measure of accessibility in accordance with the 2-Step Floating Catchment Area approach. Moreover, the proportional allocation balancing factors in @soukhovIntroducingSpatialAvailability2023 apply a singly-constraint to the accessibility measure, akin to the constraints introduced within the spatial interaction modelling framework outlined in @wilson1971.

Taken this motivation of incorporating competition in the unified framework and recent developments in methodology, this paper contends that accessibility research must reconnect with its spatial interaction origins. Particularly, we argue an important aspect of spatial interaction modelling -namely, constraining the results to match empirical observations- was never effectively reincorporated into accessibility analysis. Empirical constraints were embraced by early spatial interaction literature until the research stream diverged following the work of @hansen1959 and @wilson1971; at which point 'accessibility' and 'spatial interaction modelling' began to flow more distinctly. The application of @wilson1971's empirical constraints supported the development of various spatial interaction models that remain relevant in research and practice today [@ortuzar_2011_modelling]. However, the same cannot be traced for contemporary accessibility literature, where empirical constraints were not explicitly adopted. We argue the absence of empirical constraints or attendance proportionality constants has contributed to some of accessibility analysis' interpretability issues; for instance, the fuzziness of insights beyond simple proportional statements like 'higher-than' or 'lower-than' [@millerAccessibilityMeasurementApplication2018].

These streams of literature share common headwaters. It is by looking to the past that we believe accessibility analysis can newly wade into the future. Hence, this work's primary focus is to show that the same empirical constraints used in @wilson1971's family of spatial interaction models can be mapped onto accessibility. We begin by tracing the development of accessibility from its origins in spatial interaction, from the Newtonian gravitational expression in @ravensteinLawsMigration1889 through to the "seminal" accessibility work of @hansen1959. We then present evidence for a narrative highlighting the marked divergence between accessibility and spatial interaction modelling research after the work of @wilson1971. Next, we hark back to @wilson1971's spatial interaction models, and use it to derive a family of accessibility measures based on different types of constraints. We illustrate various members of this family with a simple numerical example and a real world data set. We then conclude by discussing the uses of these measure and their interpretation.

@@

The patterns of people's movement in space have been a subject of scientific inquiry for almost a century and a half. From as far back as Henry C. Carey's _Principles of Social Science_ [@careyPrinciplesSocialScience1858], a concern with the scientific study of human spatial interaction can be observed. It was in this work where Carey stated that "man [is] the molecule of society [and their interaction is subject to] the direct ratio of the mass and the inverse one of distance" [@mckeanManual1883, pp. 37-38]. This statement shows how investigations into human spatial interaction have often been explicitly coloured by the features of Newton's Law of Universal Gravitation, first posited in 1687's _Principia Mathematica_ and expressed as in @eq-phys-grav-prop.

@@

To be certain, the expression above, a proportionality, is one of the most famous in all of science. In brief, it states that the force of attraction $F$ between a pair of bodies $i$ and $j$ is directly _proportional_ to the product of their masses $M_i$ and $M_j$, and inversely _proportional_ to the square of the distance between them $D_{ij}$. Direct proportionality means that as the product of the masses increases, so does the force. Likewise, inverse proportionality means that as the distance increases, the force decreases. @eq-phys-grav-prop, however, does not quantify the magnitude of the force. To do so, an empirical constant is required to convert the proportionality into an equality, ensuring that values of the force $F$ in @eq-phys-grav-prop match the observed force of attraction between masses. In other words, @eq-phys-grav-prop needs to be _constrained_ using empirical data. Ultimately, the equation for the force is as seen in @eq-phys-grav, where $G$ is an empirically calibrated proportionality constant:

@@

Following Carey's _Principles_ of 1858, research into human spatial interaction continued in different contexts. In the late 1880s, Ravenstein proposed some "Laws of Migration" based on his empirical analysis of migration flows in various countries [@ravensteinLawsMigration1885; @ravensteinLawsMigration1889]. In these works, Ravenstein posited 1) a directly proportional relationship between migration flows and the size of destinations (i.e., centres of commerce and industry), and 2) an inversely proportional relationship between the size of flows and the separation between origins and destinations. As with Carey, these propositions are similar to Newton's gravitational laws. Over time, other researchers discovered similar relationships. <!--Rilley does not cite Newton, nor Carey, nor Ravenstein, but I removed this curiosity as it's not central to our point here--> For example, @reilly1929methods formulated a law of retail gravitation, expressed in terms of equal attraction to competing retail destinations. Later, Zipf proposed a $\frac{P_1P_2}{D}$ hypothesis for the case of goods movement by railways [@zipfHypothesisCaseRailway1946], intercity personal movement [@zipfHypothesisIntercityMovement1946], and information [@zipfDeterminantsCirculationInformation1946]. The $\frac{P_1P_2}{D}$ hypothesis stated that the magnitude of flows was proportional to the product of the populations of settlements, and inversely proportional to the distance between them.

@@

\noindent Where:

@@

\noindent Where $D$ is the population density over an infinitesimal area $ds$ and $r$ is the distance to $i$. In @eq-stewart-population-potential-integral, $D\cdot ds$ gives an infinitesimal count of the population, say $dm$, and so, after discretizing space, @eq-stewart-population-potential-integral can be rewritten as:
$$
V_i = \sum_j \frac{M_j}{d_{ij}}
$${#eq-stewart-population-potential-sum}

@@

Stewart's formulation of demographic force, developed in the context of what he called "social physics" [@stewartPrinciples1947], was problematic. It had issues with inconsistent mathematical notation. <!--For example, in @stewartDemographicGravitationEvidence1948, $G$ is used as a proportionality constant p. 34 and then later as _demographic energy_ on p. 53.--> More seriously though, Stewart's work was permeated by a view of humans as particles following physical laws, but tinted by unscientific ideas that were unadultered racism. For instance, he assumed that the molecular weight $\mu$ of the average American was one, but "presumably...much less than one....for an Australian aborigine" [p. 35]. Stewart's ideas about "social physics" soon fell out of favour among social scientists, but not before influencing the nascent field of accessibility research, as detailed next.

@@

From @stewartDemographicGravitationEvidence1948, we arrive to 1959 and Walter G. Hansen, whose work proved to be exceptionally influential in the accessibility literature [@hansen1959]. In his seminal paper, Hansen defined accessibility as "the potential of opportunities for interaction... a generalization of the population-over-distance relationship or _population potential_ concept developed by @stewartDemographicGravitationEvidence1948" (p. 73). As well as being a student of city and regional planning at Massachusetts Institute of Technology, Hansen was also an engineer with the Bureau of Roads, and preoccupied with the power of transportation to shape land uses in a very practical sense. @hansen1959 focused on @stewartDemographicGravitationEvidence1948's _population potential_ (expressed in @eq-stewart-population-potential-sum), leaving Stewart's other formulaic contributions and objectionable aspects of "social physics" behind. @hansen1959 recast @eq-stewart-population-potential-sum to reflect accessibility, a model of human behaviour useful to capture regularities in mobility patterns. In this equation, @hansen1959 replaced $M_j$ with _opportunities_ to derive an _opportunity potential_, or more accurately, a _potential of opportunities for interaction_ as follows:

@@

Or a contemporary rewriting of the above @eq-accessibility accounts for a variety of impedance functions beyond the inverse power $d^{-\beta}$:

@@

<!--SUMMARY: Introduce Wilson's spatial interaction model. Then discuss how it could relate to gravity analogy -- but is tactically derived using entropy-maximisation, obtaining the interaction variable as a statistical average. Then introduce the constraints and cases. -->

@@

Notice how, in this model, the proxy for the mass at the origin $W_i^{(1)}$ is replaced by the actual mass, as measured by the trips produced $O_i$. Also, there is no longer a single system-wide proportionality constant, but rather a balancing factor specific to each origin, namely $A_i$, which Wilson found to be as follows:

@@

Notice how now the proxy for the mass at the destination $W_j^{(2)}$ is replaced by the actual mass, as measured by the trips attracted $D_j$. As before, destination-specific balancing factors $B_j$ were derived by Wilson as:

@@

In this model, both proxies for the masses are replaced with the known masses, that is, the trips produced by origin and the trips attracted by destination. And, now the balancing factors are:

@@

The work of @hansen1959 and @wilson1971 responded to important developments in the Anglo-American world at the time, in particular a need "to meet the dictates and needs of public policy for strategic land use and transportation planning" [@battyChronicleScientificPlanning1994]. Said dictates and needs were far from trivial. In the US alone, the Federal-Aid Highway Act of 1956 authorized the creation of the US Interstate Highway System, with a budget of more than one hundred billion dollars [@weinerUrbanTransportationPlanning2016; @mdotMnDOTJoins2007]. Spatial interaction modelling, with its ability to quantify trips, was incorporated into institutional modelling practices meant to "predict and provide", i.e., predict travel demand and supply transportation infrastructure [@kovatch1971modeling; @weinerUrbanTransportationPlanning2016]. Accessibility at the time did not quite have that power: it did not quantify trips, but rather something more elusive: the less tangible "potential for interaction". In this way, where spatial interaction modelling became a key element of transportation planning practice, accessibility remained a somewhat more academic pursuit, and the two streams of literature only rarely connected.

@@

To illustrate this point, we conducted a bibliographic analysis of the literature that cites @hansen1959, @wilson1971 or both. We retrieved all relevant documents using the Web of Science "Cited References" functionality, and the digital object identifiers of @hansen1959 and @wilson1971. As a result of this search we identified `r n_h |> prettyNum(big.mark = ",")` documents that cite @hansen1959, `r n_w  |> prettyNum(big.mark = ",")` documents that cite @wilson1971, and `r n_b |> prettyNum(big.mark = ",")` that cite both. The earliest document in this corpus dates to `r min(corpus$PY)` and the most recent is from `r max(corpus$PY)`. The number of documents per year appears in @fig-docs-per-year, where we see the frequency of documents over a span of almost fifty years. In particular, we notice the remarkable growth in the number of papers that cite @hansen1959 compared to those that cite @wilson1971 or both.

@@

After compiling this corpus of documents, we used the {bibliometrix} package [@bibliometrix2017] to create a bibliographical coupling matrix. In this matrix, two documents are coupled if they share at least one common reference, and the strength of the coupling increases with the number of references that the documents have in common. @fig-Fig1 presents the results. In this figure, each symbol represents a document and their distance on the plot represents the strength of bibliographical coupling: more strongly coupled documents suggests those that share more items in their lists of references are plotted closer to each other.

@@

As we argued in the preceding sections, the accessibility and spatial interaction modelling literature streams tended to evolve with relatively little contact. This may explain the failure of the constraints/proportionality constant(s) of spatial interaction models to cross over to accessibility analysis. This is intriguing, considering that Wilson himself made an effort to connect his developments in spatial interaction modelling to accessibility, noting the similarity between the denominator in @eq-production-constrained-balancing-factor and Hansen's accessibility indicator. In fact, Wilson noted that Hansen's accessibility was the inverse of $A_i$, which offered a potential interpretation for the balancing factor (p. 10):

@@

$$
V^T_i = k\sum_j W^{(2)}_jf(c_{ij}) = \sum_j W^{(2)}_j\frac{O*f(c_{ij})}{\sum_i\sum_j V^0_{ij}} = \sum_j W^{(2)}_j\frac{O*f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})} 
$$

@@

$$
\kappa = \frac{O*f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
$$

@@

In the case of SIM calibrated for accessibility, the doubly-constrained case has more limited applications than the singly-constrained case because the double constraints (both marginals) are so specific. In this case, the units of the population and opportunities must be able to be understood as the same. E.g., the number of population at $i$ that seek opportunities at each $j$ and the number of opportunities at $j$ that are sought by population at each $i$ must both be known to match.
